A semantics of core AUTOSAR

### 2016-09-01 paper/paper.tex
Tiny fixes.

Closely related to the interleaving of timing events, and of interest even if only distinct runnables are being triggered, is the notion of initial timer offsets. That is, at what time will each timer tick for the first time --- immediately at startup or at some other (later) time? The AUTOSAR specification only introduces a concept of timing offsets in relation to the operating system tasks that are expected to implement runnables; the Software Component Template itself is totally silent on the issue \cite{AR:SWC}. Lacking an abstract definition of runnable offsets, we have chosen to simply fire all timers simultaneously at $0$ seconds from startup (hence the value $0$ in the initial {\bf timer} term definition in Section~\ref{sec:Sem}). It is not hard to see that other values may be equally reasonable, though, like
$$
	\Ttimer {i.r} {t_p} {t_p}
$$
which lets each timer wait one full cycle before triggering for the first time. However, should we really want to take the assumed task offsets into account, the initial timer state would rather be
$$
	\Ttimer {i.r} {t_p} {t_o}
$$
for some non-negative $t_o$. One might even want to add the condition $t_o < t_p$, but it must be emphasized that all these initialization options are speculative --- the AUTOSAR Software Component Template must be clarified before a sufficiently abstract model of timed behavior can be obtained.

@@

The formal semantics may also provide a foundation for tools that analyze or transform AUTOSAR models. One such example is our AUTOSAR simulator \cite{ARSim}, which interprets AUTOSAR models entirely on basis of the transition system, utilizing truly random scheduling and a maximum progress assumption to select among available transitions at each step. The benefit of this simulator is its ability to dynamically execute AUTOSAR Software Component models as they are, without prior commitment to a particular platform, or translation into a low level representation. A further avenue of exploration concerns the potential of proving properties of AUTOSAR systems on the basis of their semantics. To prepare for this opportunity, we have defined the transition system used in this paper as a Horn clause file readable by common theorem provers, and, in fact, Appendix~\ref{sec:Rules} is automatically produced through a simple formatting translation of this file \cite{AR.eprover}.

>>>>

Closely related to the interleaving of timing events, and of interest even if only distinct runnables are being triggered, is the notion of initial timer offsets. That is, at what time will each timer tick for the first time --- immediately at startup or at some other (later) time? The AUTOSAR specification only introduces a concept of timing offsets in relation to the operating system tasks that are expected to implement runnables; the Software Component Template itself is totally silent on the issue \cite{AR:SWC}. Lacking an abstract definition of runnable offsets, we have chosen to simply fire all timers simultaneously at $0$ seconds from startup (hence the value $0$ in the initial {\bf timer} term definition in Section~\ref{sec:Sem}). It is not hard to see that other values may be equally reasonable, though, like
$$
	\Ttimer {i.r} {t_p} {t_p}
$$
which lets each timer wait one full cycle before triggering for the first time. However, should we really want to take the assumed task offsets into account, the initial timer state would rather be
$$
	\Ttimer {i.r} {t_p} {t_o}
$$
for some non-negative $t_o$. One might even want to add the condition $t_o \leq t_p$, but it must be emphasized that all these initialization options are speculative --- the AUTOSAR Software Component Template must be clarified before a sufficiently abstract model of timed behavior can be obtained.

@@

The formal semantics may furthermore provide a foundation for tools that analyze or transform AUTOSAR models. One such example is our AUTOSAR simulator \cite{ARSim}, which interprets AUTOSAR models entirely on basis of the transition system, utilizing truly random scheduling and a maximum progress assumption to select among available transitions at each step. The benefit of this simulator is its ability to dynamically execute AUTOSAR Software Component models as they are, without prior commitment to a particular platform or translation into a low level representation. A further avenue of exploration concerns the potential of proving properties of AUTOSAR systems on the basis of their semantics. To prepare for this opportunity, we have defined the transition system used in this paper as a Horn clause file readable by common theorem provers, and, in fact, Appendix~\ref{sec:Rules} is automatically produced through a simple formatting translation of this file \cite{AR.eprover}.
